\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-07-01 paper/main.tex
.

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. On one hand, molecular information can be used
to directly determine whether two molecules (such as a drug and
protein) have the ability to interact, whereas information about
phenotypic consequences of a drug (drug effects) or disruption of a
protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular features will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Moreover, as the protein--protein 
interaction relation is not transitive (protein $P_1$ interacting with $P_2$, 
and $P_2$ relating with $P_3$ does not imply $P_1$ interacting
or influencing $P_3$), we can only infer from the direct neighborhood of each protein within the PPI graph. 
Our results (Table
\ref{tab:Results}) confirm the first hypothesis and demonstrate that
molecular features do not benefit from including the interaction
network whereas the indirect, top-down features benefit from the
propagation over the interaction network. While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction. Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.

@@

>>>>

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. On one hand, molecular information
can be used to directly determine whether two molecules (such as a
drug and protein) have the ability to interact, whereas information
about phenotypic consequences of a drug (drug effects) or disruption
of a protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular features will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Moreover, as the
protein--protein interaction relation is not transitive (protein $P_1$
interacting with $P_2$, and $P_2$ relating with $P_3$ does not imply
$P_1$ interacting with $P_3$), we can only infer from the direct
neighborhood of each protein within the PPI graph. Our results (Table
\ref{tab:Results}) confirm the first hypothesis and demonstrate that
molecular features do not benefit from including the interaction
network whereas the indirect, top-down features benefit from the
propagation over the interaction network.

@@

\name{} introduces a number of novel

While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction. Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.
